### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

The automaton $A$ is extended by a set $X=X_1\times ... \times
X_n$, where $X_k$, $k=1,...,n$, are a finite nonempty subsets of
$\mathbb{Z}$. The set $X$ is the domain of definition of an
$n$-tuple of integer variables $x=(x_1,...,x_n)$ with initial
value $x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple
of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
(g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
$(g_{\tau_k},a_{\tau_k})$ is associated to the corresponding
transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
boolean formula (a guard formula) over $X$ i.e. $g_{\tau_k}:X
\rightarrow \{0,1\}$ saying if the corresponding transition
$\tau_k$ is enabled or not. $a_{\tau_k}$ is an updating (action)
function from $X$ to $X$ of the integer variables
$x=(x_1,...,x_n)$.

@@

\subsection{Synchronous composition of EFA's}
The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
exists if and only if the automata are consistent. The consistency
condition requires that the action functions triggered by a common
enabled event $\sigma\in \Sigma_{A}\cap \Sigma_{B}$ in the two
automata update common variables to the same values, i.e. if the
$i^{ith}$ variable in $x^{A}$ and the $j^{ith}$ variable in
$x^{B}$ are the same variable and $g_{q^A \sigma}(x^{A})\wedge
g_{q^B \sigma}(x^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
$(a_{q^A \sigma}(x^{A}))_i = (a_{q^B \sigma}(x^{B}))_j$.

@@

 Let $z^{AB}$ denote the common variables of $x^{A}$ and $x^{B}$, and let $z^A$
and $z^B$ be the distinct variables of $x^{A}$ and $x^{B}$,
respectively. We define the integer variables $z$ of the
synchronous product of $A$ and $B$ as the vector $z=(z^A, z^{AB},
z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(z)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
\sigma\in\Sigma^B\setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
write equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(z)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
To define the action functions for $A||B$ we look at the
restrictions of the updating functions $a^A_\delta$ and
$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
are defined as (if $x^{A}$ and $x^{B}$ have no common variables
then $z=(x^{A},x^{B})$)
\begin{eqnarray}\label{action_sync_1}
a_{q \sigma}(z)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A \sigma}(x^{A}), z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
\big(z^{A}, a^B_{q^B \sigma}(x^{B})\big)\quad\quad\quad\sigma\in
\Sigma^B \setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
variables then equation (\ref{action_sync_1}), that defines the
action functions of the product, simplifies to
\begin{eqnarray}
a_{q \sigma}(z)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

@@

\begin{theorem}
   Full synchronous composition of EFAs is an associative
   operation.
\end{theorem}

\noindent \textbf{Proof}: To prove the associativity we
considering three automata $A, A'$ and $A''$, and show that
$(A\|A')\|A'' = A\|(A'\|A'')$. First, we assume that the three
automata have no common variables.
\newline\newline
\noindent The 8-tuple representing $(A\|A')\|A''$ is
\begin{center}
\begin{tabular}{ll}
  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
  & $\delta^{(A\|A')\|A''},$\\
  & $\big((q_i, q'_i) , q''_i\big),$\\
  & $(Q_m \times Q_m') \times Q_m'',$\\
  & $\big[[X, X'], X'']$\\
  & $G^{(A\|A')\|A''},$\\
  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
\end{tabular}
\end{center}

The variables for the product are $X^c=(X_m, X'_m , X''_m)$ and
the number of variables are $n=|X^c|$. For the guard functions
$g^{(A\|A')\|A''} : {\mathbb{N}}^{n} \rightarrow \{\mathrm{true},
\mathrm{false}\}$ of the composite automaton we have by definition

klkljkll

>>>>

The automaton $A$ is extended by a set $X=X_1\times ... \times
X_n$, where $X_k$, $k=1,...,n$, are finite nonempty subsets of
$\mathbb{Z}$. The set $X$ is the domain of definition of an
$n$-tuple of integer variables $x=(x_1,...,x_n)$ with initial
value $x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple
of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
(g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
$(g_{\tau_k},a_{\tau_k})$ is associated to the corresponding
transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
boolean formula (a guard formula) over $X$ i.e. $g_{\tau_k}:X
\rightarrow \{0,1\}$ saying if the corresponding transition
$\tau_k$ is enabled or not. $a_{\tau_k}$ is an updating (action)
function from $X$ to $X$ of the integer variables
$x=(x_1,...,x_n)$.

@@

\subsection{Synchronous composition of EFA's}

Formally we define the synchronous product $A||B$ of two EFA's
\begin{eqnarray}
A=\langle Q^A, \Sigma^A, \delta^A, q^A_i, Q^A_m, X^{A},
G^A,x^A_0,X^A_m \rangle
\end{eqnarray} and \begin{eqnarray} B=\langle Q^B, \Sigma^B, \delta^B, q^B_i,
Q^B_m, X^{B}, G^B,x^B_0,X^B_m \rangle
\end{eqnarray} as
\begin{eqnarray}
A||B&=&\langle Q^A\times Q^B, \Sigma^A \cup \Sigma^B ,
\delta^{A||B}, (q^A_i,q^B_i) ,\nonumber\\
&& Q^A_m\times Q^B_m, X^A \times X^B, G^{A||B},(x^A_0,x^B_0),
X^A_m \times X^B_m \rangle.
\end{eqnarray}
To define the guard functions, the action functions and the
transition function of the product we make use of the
corresponding functions in $A$ and $B$. To denote the elements of
$A||B$ we remove unnecessary  superscripts and write $A||B=\langle
Q, \Sigma, \delta, q_i, Q_m, X, G ,x_0, X_m \rangle$. The guard
functions $g_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(x)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A})\quad\quad\quad\quad\quad\quad\quad\sigma\in \Sigma^A\setminus\Sigma^B\\
g_{q^B\sigma}(x^{B})\quad\quad\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. Doing this, we can write
equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(x)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
The updating functions $a_{q \sigma}(x)$ are defined as
\begin{eqnarray}\label{action_sync_1}
a_{q \sigma}(x)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A\sigma}(x^{A}),x^{B}\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^A\setminus\Sigma^B\\
\big(x^{A},a^B_{q^B\sigma}(x^{B})\big)\quad\quad\quad\quad\quad\sigma\in\Sigma^B\setminus\Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. With this convention equation
(\ref{action_sync_1}) simplifies to
\begin{eqnarray}
a_{q \sigma}(x)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

 The transition function
$\delta(q,\sigma,x)$ is defined from $\delta^A$ and $\delta^B$ as
\begin{eqnarray}
\delta(q,\sigma,x)=\left\{
\begin{array}{ll}
\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(x)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(x)=1\\
\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(x)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(x)=1\\
\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(x)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(x)=1\\
\textrm{
undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
g_{q\sigma}(x)=0
\end{array}\right.
\end{eqnarray}

\begin{theorem}
   Full synchronous composition of EFA's is an associative
   operation.
\end{theorem}

\noindent \textbf{Proof}: To prove the associativity we
considering three automata $A, A'$ and $A''$, and show that
$(A\|A')\|A'' = A\|(A'\|A'')$. Here, we assume that the three
automata have no common variables.
\newline\newline
\noindent The 8-tuple representing $(A\|A')\|A''$ is
\begin{center}
\begin{tabular}{ll}
  $(A\|A')\|A'' =$ & $\Big \langle (Q \times Q') \times Q'',$\\
  & $ \big(\Sigma \bigcup \Sigma' \big) \bigcup \Sigma'',$\\
  & $\delta^{(A\|A')\|A''},$\\
  & $\big((q_i, q'_i) , q''_i\big),$\\
  & $(Q_m \times Q_m') \times Q_m'',$\\
  & $\big[[X, X'], X'']$\\
  & $G^{(A\|A')\|A''},$\\
  & $\big[[X_m, X'_m], X''_m \big] \Big \rangle$
\end{tabular}
\end{center}

For the guard functions $g^{(A\|A')\|A''} : X^c \rightarrow
\{\mathrm{true}, \mathrm{false}\}$ of the composite automaton we
have by definition

\subsection{Synchronous composition of EFA's with common variables}

The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
with common variables exists if and only if the automata are
consistent. The consistency condition requires that the action
functions triggered by a common enabled event $\sigma\in
\Sigma_{A}\cap \Sigma_{B}$ in the two automata update common
variables to the same values, i.e. if the $i^{ith}$ variable in
$x^{A}$ and the $j^{ith}$ variable in $x^{B}$ are the same
variable and $g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})=1$
where $(q^A,q^B)\in Q^A\times Q^B$ then $(a_{q^A \sigma}(x^{A}))_i
= (a_{q^B \sigma}(x^{B}))_j$.

@@

 Let $z^{AB}$ denote the common variables of $x^{A}$ and $x^{B}$, and let $z^A$
and $z^B$ be the distinct variables of $x^{A}$ and $x^{B}$,
respectively. We define the integer variables $z$ of the
synchronous product of $A$ and $B$ as the vector $z=(z^A, z^{AB},
z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
\begin{eqnarray}\label{guard_sync_2}
g_{q \sigma}(z)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
\sigma\in\Sigma^B\setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
write equation (\ref{guard_sync_2}) as
\begin{eqnarray}
g_{q \sigma}(z)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
To define the action functions for $A||B$ we look at the
restrictions of the updating functions $a^A_\delta$ and
$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
are defined as (if $x^{A}$ and $x^{B}$ have no common variables
then $z=(x^{A},x^{B})$)
\begin{eqnarray}\label{action_sync_2}
a_{q \sigma}(z)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A \sigma}(x^{A}), z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
\big(z^{A}, a^B_{q^B \sigma}(x^{B})\big)\quad\quad\quad\sigma\in
\Sigma^B \setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
variables then equation (\ref{action_sync_2}), that defines the
action functions of the product, simplifies to
\begin{eqnarray}
a_{q \sigma}(z)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

@@
